Using VDMS to Search 100M Images: A Piece of Cake

### 2020-02-03 yfcc100m/paper/vldb2020/eval.tex
Update FeatureVector results

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities or visual objects.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
These vectors can be labeled and classified to build search indexes,
and there are many in-memory libraries that are designed for
this task~\cite{flann, faiss}.
Using the VDMS API, users can manage feature vector indexes,
query previously inserted elements (images),
run a k-nearest neighbor search (knn) and express relationships
between existing images or feature vectors and
the newly inserted feature vectors.
By natively supporting feature vectors and knn,
VDMS allows out-of-the-box classification
functionalities for many applications.

For this work, we wanted to

% Figure~\ref{fig:features} shows an example of how this functionality
% can be integrated in a medical imaging analytics pipeline:
% (1) A feature vector is extracted from an bounding
% box in a brain image and labeled;
% (2) the feature vector is inserted together
% with all the associated metadata (type of cancer, for example)
% and a link to the original image.
% (3) A new feature vector is extracted from a new image,
% (4) a query to VDMS is issued to classify that feature vector based on the
% indexed features in VDMS, and
% (5) VDMS can respond to the query with the label
% associated to that feature vector.
% Note that, even if the example is based on the medical images used for the
% performance evaluation, this methodology is applicable in many other contexts
% and use cases, such as face detection and matching.
% More details about the JSON API for this functionality can be found on
% our Github wiki page.

@@

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/features_alternatives}
\caption{Feature Vector Evaluation}
\label{fig:features_eval}
\end{figure*}

@@

>>>>

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities or visual objects.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as "descriptors" or "visual descriptors".
We use these terms interchangeably.
These descriptors can be labeled and classified to build search indexes,
and there are many in-memory libraries that are designed for
this task~\cite{flann, faiss}.
Using the VDMS API, users can manage feature vector indexes,
query previously inserted elements (images),
run a k-nearest neighbor search (knn) and express relationships
between existing images or descriptors and
the newly inserted descriptors.
By natively supporting descriptors and knn,
VDMS allows out-of-the-box classification
functionalities for many applications.

For this work, and as part of a comprehensive image search implementation,
we have used 4096-dimensional descriptors extracted from every image
(and first frame of every video) and created a collection of these feature
vectors in VDMS to be able to perform similarity search (i.e., find
images that are "similar" to an query (input) image).
"Similarity" in this particular case is defined as closeness
in a an 4096-dimensional space using euclidean distance as the metric.

EXPLAIN HERE DIFFERENT ACCURACY/EXECUTIME TIME TRADEOFF.

The process of loading descriptors in VDMS is simple.
First, the user has to create a DescriptorsSet, using a single command.
At creation of the DescriptorSet, the dimensionality of the descriptors
is specified, together with the desired indexing method and the desired metric
for computing distances (Euclidean Distance, L2, or Inner Product, IP).
Once the DescriptorSet is created, descriptors can be inserted to the set.
After the descriptors are inserted, similarity search can be performed.
Note that descriptors can be inserted later on, and any search will reflect
the existence of all descriptors in the set.

@@

Figure \ref{fig:similarity} shows 3 examples of a query image (on the left),
and images returned as "similar" by VDMS.
The query input is a descriptor generated after a query image. The "query"
descriptor is sent to VDMS as part of the query, and VDMS use that descriptor
to find similar one, and retrieve the images associated with those "similar"
descriptors. We used this as an example and as a visual validation of the
functionality and applicability in this particular dataset, but we also
provide an analytical approach to accuracy and trade-offs in our system.
It is important to note that the accuracy of the results is entirely tied
to the quality of the descriptors chosen by the applications.
VDMS is completely agnostic to information contained within the descriptor,
and simply offers the interface to store and index them, but the quality
of the similarity result will be tied to the quality of descriptor extraction
that the application is using.

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/features_alternatives}
\caption{Feature Vector Evaluation: Trade-off between query execution speed
and accuracy of the results, using ground-truth data for computing accuracy.
For this evaluation, we query the 10 closest neighbors (k = 10), and compute
accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
results that is present within the top 10 computed neighbors).
We average the query execution time and accuracy for 100 queries (nq = 100).}
\label{fig:features_eval}
\end{figure*}

As mentioned before, VDMS provides different levels of customization of the
indexes created for a descriptor set, that includes the indexing techniques
and the metric for similarity.
These different indexing techniques comes with different trade-offs in terms
of speed of search and accuracy.
VDMS aims to provide functionality that is agnostics to application-specific
techniques, enabling functionality that is general to visual data processing
applications.
Figure \ref{fig:features_eval} shows an analysis at the different indexing
techniques provided by VDMS and its trade-off between accuracy and query
execution speed, for a single threaded client.
For this evaluation, we query the 10 closest neighbors (k = 10), and compute
accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
results that is present within the top 10 computed neighbors).
We average the query execution time and accuracy for 100 queries (nq = 100).
The "flat" index (red line) implements exact search and
represents ground-truth, which explain why the accuracy is always 100\% on the
right plot. The other indexes implement "approximate search", which trade-offs
between accuracy and speed of search~\cite{flann, faiss}.
We have also tried the "ivfflat" index (inverted file index), as well as
LSH-based indexes using a different number of bits per descriptor.
\footnote{https://github.com/facebookresearch/faiss/wiki/Faiss-indexes}.
Results show how "ivfflat" is the fastest option by comes at the trade-off
of about 30\% lost in accuracy, while simple brute-force search
is among the slowest options at the expenses of 100\% accuracy,
meaning exact search.

@@

Another important trade-off to be made is w.r.t to space efficiency: Set of
descriptors can grow very large and hard to manage.
In this particular case, 4096-dimensional descriptors for 100M elements
translates into 1TB of data, only in raw floating-point data alone (without
accounting for any metadata or indexes associated with it).
This component is very important on the overall analysis because a large set
of feature vectors may not fit in memory and thus cause a pressure on the IO
system while retrieving descriptors for computing distance, that severely impact
the overall query execution time.
Because of this, when the set of descriptors grows significantly large,
it may be worth trading off accuracy to speed and space.
Figure \ref{fig:features_size_does_matter} show the different indexes and
their size in disk. These indexes already contain all the descriptors (or
a quantized version of them), and can be loaded in memory directly when it fits.
Note how, because of quantization of the descriptors,
LSH provides a significantly lower space foot print, which can be a great option
for large collections of descriptors when accuracy is not a main factor.

### 2020-02-03 yfcc100m/paper/vldb2020/intro.tex
Update FeatureVector results

In this paper, we also expand on the Video and Feature Vector capabilities of
VDMS, which are part of the latest additions to the system.
We analyse different functionalities and trade-offs for this type of data,
in combination with metadata filtering. To the best of out knowledge, this
set of functionalities are unique to VDMS and we were not able to find a system
with similar functionality.

>>>>

In this paper, we also expand on the Video and Feature Vector capabilities of
VDMS, which are part of the latest additions to the system.
We analyze different functionalities and trade-offs for this type of data,
in combination with metadata filtering. To the best of out knowledge, this
set of functionalities are unique to VDMS and we were not able to find a system
with similar functionality.

We show how VDMS can be used as the single and centralize point for data
management and data access even when having multiple modalities of data:
Metadata, Image, Videos, and Feature Vectors, all part of a comprehensive
dataset that is used for multiple use-cases.
